Staged Abstract Interpreters

### 2019-03-29 paper/70_eval.tex
related work

We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04LTS
(kernel 4.4.0). All of our evaluation were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores and 192 threads in total, the abstract interpreters only use one thread.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

>>>>

We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
(kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
cores (192 threads) in total, the abstract interpreters are single-threaded.
To minimize the effect of warming-up from the JVM JIT compiler, all the
experiments are executed for 20 times and we report the statistical median value
of the running times (in seconds).

### 2019-03-29 paper/80_related.tex
related work

\paragraph{Optimizing Static Analysis Through Specialization}
The idea in this paper is closely inspired by abstract compilation
\cite{Boucher:1996:ACN:647473.727587}.
\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
techniques as an efficient implementation of the monovariant flow analysis
(\textit{0}-CFA) for programs written in continuation-passing style. The key
idea is to remove the interpretation overhead on traversing the syntax tree by
partial evaluation. Specifically, they proposed two similar kinds of abstract
compilation techniques. The first one is to generate specialized analysis as a
textual program, which then can be loaded and executed by \texttt{eval} or other
similar mechanisms. The second one is to use closures, i.e., functions that
remember their environments, as a representation of specialized analysis. As we
show in the case study, compiling the analysis generates higer-order functions
on-the-fly with respect to the analyzed program, then the generated closure can
be applied immediately in the higher-order host language.

\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
generation to optimize small-step abstract interpreter in state-transition style
\cite{Johnson:2013:OAA:2500365.2500604}. The analyzed program is firstly
compiled to a intermediate representation called "abstract bytecode", which are
actually higher-order functions, and then be executed on a abstract abstract
machine for that IR.

\citet{damian1999partial} provides a formal treatment to abstract compilation
and Shiver's CFA, as well as proofs to establish correctness of the certified
specialized analyzer \cite{damian1999partial}. \citeauthor{amtoft1999partial}
applied partial evaluation for constraint-based control flow analysis
\cite{amtoft1999partial}. Splitting an analysis into multiple stages is also
studied for analysis other than for control-flow, though the formulation may very
different. For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to
flow-sensitive pointer analysis \cite{DBLP:conf/cgo/HardekopfL11}. The first
stage is to analyze the program code to obtain a sparse representation, and then
the second stage conducts the flow-sensitive analysis based on the first one.

\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domains
through online decomposition \cite{DBLP:conf/popl/SinghPV17,
Singh:2017:PCD:3177123.3158143}. The idea is to decompose a large polyhedra into
several smaller one; one abstract transformer usually only depends a limit
number of variables, then we just need to compute the resulting polyhera on
those smaller polyhedras which contain these related variables. The decomposition
can be considered as a specialization of abstract domain with respect to the
static program structure.

>>>>

\paragraph{Optimizing Static Analysis Through Specialization}
The underlying idea in this paper is closely related to abstract compilation (AC)
\cite{Boucher:1996:ACN:647473.727587}: removing the interpretation overhead on
traversing the syntax tree by specialization. Specifically, the residual program
of AC can be either textual or closures. However, as we studied in Section
\ref{cs_ac}, with the perspective from high-performance DSLs and monadic
interpreters, we make it more general, extensible and easy.
\citet{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
generation to optimize small-step abstract interpreter in state-transition
style. The analyzed program is firstly compiled to a IR called "abstract
bytecode", which are actually higher-order functions and will be executed later
on a abstract abstract machine for that IR. \citet{damian1999partial} provides a
formal treatment to abstract compilation and Shiver's CFA, as well as proofs to
establish correctness of the certified specialized analyzer.
\citet{amtoft1999partial} applied partial evaluation for constraint-based
control flow analysis. Splitting an analysis into multiple stages is also
studied for analysis other than for control-flow, though the formulation may
very different. For example, \citet{DBLP:conf/cgo/HardekopfL11} apply staging to
flow-sensitive pointer analysis. The first stage is to analyze the program code
to obtain a sparse representation, and then the second stage conducts the
flow-sensitive analysis based on the first one.
The abstract domains can be also specialized w.r.t. the static program
structure, e.g., decomposing polyhedras \cite{DBLP:conf/popl/SinghPV17,
Singh:2017:PCD:3177123.3158143} to smaller ones such that the abstract
transformers can compute using less time and space.
